### 2018-08-20 document/manuscript-biorxiv.tex
Initial set up of paper template

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in,marginparwidth=2in]{geometry}

% use Unicode characters - try changing the option if you run into troubles with special characters (e.g. umlauts)
\usepackage[utf8]{inputenc}

% clean citations
\usepackage{cite}

% hyperref makes references clicky. use \url{www.example.com} or \href{www.example.com}{description} to add a clicky url
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% improves typesetting in LaTeX
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% text layout - change as needed
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Remove % for double line spacing
%\usepackage{setspace} 
%\doublespacing

% use adjustwidth environment to exceed text width (see examples in text)
\usepackage{changepage}

% adjust caption style
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,singlelinecheck=off]{caption}

% remove brackets from references
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% headrule, footrule and page numbers
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}

% use \textcolor{color}{text} for colored text (e.g. highlight to-do areas)
\usepackage{color}

% define custom colors (this one is for figure captions)
\definecolor{Gray}{gray}{.25}

% this is required to include graphics
\usepackage{graphicx}

% use if you want to put caption to the side of the figure - see example in text
\usepackage{sidecap}

% use for have text wrap around figures
\usepackage{wrapfig}
\usepackage[pscoord]{eso-pic}
\usepackage[fulladjust]{marginnote}
\reversemarginpar

% document begins here
\begin{document}
\vspace*{0.35in}

% title goes here:
\begin{flushleft}
{\Large
\textbf\newline{\LaTeX Template for Manuscripts with Embedded Figures.}
}
\newline
% authors go here:
\\
Author 1\textsuperscript{1},
Author 2\textsuperscript{2},
Author 3\textsuperscript{1},
Author 4\textsuperscript{1},
Author 5\textsuperscript{2},
Author 6\textsuperscript{2},
Author 7\textsuperscript{1,*}
\\
\bigskip
\bf{1} Affiliation A
\\
\bf{2} Affiliation B
\\
\bigskip
* correseponding@author.mail

\end{flushleft}

\section*{Abstract}
Fitting multiple figures into very tight manuscripts while keeping it pleasant to read is challenging. Therefore figures are often simply attached to the very end of a manuscript file. While easier for the authors, this practice is inconvenient for readers. This \LaTeX template shows how to generate a compiled PDF with figures embedded into the text. It provides several examples of how to embed figures or tables directly into the text thus giving you a range of options from which you should choose the one best suited for your manuscript. Check out Schlegel et al., (2016) as example of use \cite{Schlegel2016}.

% now start line numbers
\linenumbers

\input manuscript

% the * after section prevents numbering

\section*{Acknowledgments}
We thank KENET for providing us with an ample environment for our hackathon.

\nolinenumbers

%This is where your bibliography is generated. Make sure that your .bib file is actually called library.bib
\bibliography{library}

%This defines the bibliographies style. Search online for a list of available styles.
\bibliographystyle{abbrv}

\end{document}

### 2018-08-21 push.sh
automate pdf push

@@ -1,0 +1,21 @@
+#!/bin/sh
+
+setup_git() {
+  git config --global user.email "travis@travis-ci.org"
+  git config --global user.name "Travis CI"
+}
+
+commit_paper_files() {
+  git add ./documents/*.pdf
+  git add ./documents/*.tex
+  git commit --message "Travis build: $TRAVIS_BUILD_NUMBER"
+}
+
+upload_files() {
+  git remote add origin https://${GIT_TOKEN}@github.com/MVSE-outreach/resources.git > /dev/null 2>&1
+  git push --quiet --set-upstream origin master
+}
+
+setup_git
+commit_paper_files
+upload_files

### 2018-08-21 document/push.sh
Update push.sh

@@ -6,8 +6,8 @@ setup_git() {
 }
 
 commit_paper_files() {
-  git add ./documents/*.pdf
-  git add ./documents/*.tex
+  git add *.pdf
+  git add *.tex
   git commit --message "Travis build: $TRAVIS_BUILD_NUMBER"
 }
 

### 2018-08-21 document/push.sh
Update push.sh

@@ -15,7 +15,6 @@ upload_files() {
   git remote add origin https://${GIT_TOKEN}@github.com/MVSE-outreach/resources.git > /dev/null 2>&1
   git push --quiet --set-upstream origin master
 }
-
 setup_git
 commit_paper_files
 upload_files

### 2018-08-21 document/push.sh
Update push.sh

@@ -12,7 +12,7 @@ commit_paper_files() {
 }
 
 upload_files() {
-  git remote add origin https://${GIT_TOKEN}@github.com/MVSE-outreach/resources.git > /dev/null 2>&1
+  git remote add origin https://${GIT_TOKEN}@github.com/bioinfonet/hack-paper.git > /dev/null 2>&1
   git push --quiet --set-upstream origin master
 }
 setup_git

### 2018-08-21 document/push.sh
Update push.sh

@@ -1,8 +1,8 @@
 #!/bin/sh
 
 setup_git() {
-  git config --global user.email "travis@travis-ci.org"
-  git config --global user.name "Travis CI"
+  git config --global user.email "calebkibet88@gmail.com"
+  git config --global user.name "Caleb Kibet"
 }
 
 commit_paper_files() {

### 2018-10-11 document/push.sh
Change on add remote push.sh

@@ -12,7 +12,7 @@ commit_paper_files() {
 }
 
 upload_files() {
-  git remote add origin https://${GIT_TOKEN}@github.com/bioinfonet/hack-paper.git > /dev/null 2>&1
+  git remote add origin https://bioinfonet:${GIT_TOKEN}@github.com/bioinfonet/hack-paper.git > /dev/null 2>&1
   git push --quiet --set-upstream origin master
 }
 setup_git

### 2018-10-11 document/push.sh
remove empty line

@@ -12,7 +12,7 @@ commit_paper_files() {
 }
 
 upload_files() {
-  git remote add origin https://bioinfonet:${GIT_TOKEN}@github.com/bioinfonet/hack-paper.git > /dev/null 2>&1
+  git remote add origin https://BioinfoNet:${GIT_TOKEN}@github.com/BioinfoNet/hack-paper.git > /dev/null 2>&1
   git push --quiet --set-upstream origin master
 }
 setup_git

### 2018-10-11 document/push.sh
format .travis.yml

@@ -12,7 +12,7 @@ commit_paper_files() {
 }
 
 upload_files() {
-  git remote add origin https://BioinfoNet:${GIT_TOKEN}@github.com/BioinfoNet/hack-paper.git > /dev/null 2>&1
+  git remote add origin https://BioinfoNet:${GITHUB_TOKEN}@github.com/BioinfoNet/hack-paper.git > /dev/null 2>&1
   git push --quiet --set-upstream origin master
 }
 setup_git

### 2018-10-11 document/push.sh
change push.sh github-token

@@ -12,7 +12,7 @@ commit_paper_files() {
 }
 
 upload_files() {
-  git remote add origin https://BioinfoNet:${GITHUB_TOKEN}@github.com/BioinfoNet/hack-paper.git > /dev/null 2>&1
+  git remote add origin https://BioinfoNet:${github-token}@github.com/BioinfoNet/hack-paper.git > /dev/null 2>&1
   git push --quiet --set-upstream origin master
 }
 setup_git

### 2018-10-11 document/push.sh
add username and token

@@ -3,6 +3,7 @@
 setup_git() {
   git config --global user.email "calebkibet88@gmail.com"
   git config --global user.name "Caleb Kibet"
+  
 }
 
 commit_paper_files() {
@@ -12,7 +13,7 @@ commit_paper_files() {
 }
 
 upload_files() {
-  git remote add origin https://BioinfoNet:${github-token}@github.com/BioinfoNet/hack-paper.git > /dev/null 2>&1
+  git remote add origin https://kipkurui:${github-token}@github.com/BioinfoNet/hack-paper.git > /dev/null 2>&1
   git push --quiet --set-upstream origin master
 }
 setup_git

### 2018-10-11 document/push.sh
add token to git push

@@ -13,8 +13,7 @@ commit_paper_files() {
 }
 
 upload_files() {
-  git remote add origin https://kipkurui:${github-token}@github.com/BioinfoNet/hack-paper.git > /dev/null 2>&1
-  git push --quiet --set-upstream origin master
+  git push https://kipkurui:${GITHUB_TOKEN}@github.com/BioinfoNet/hack-paper.git --quiet --all
 }
 setup_git
 commit_paper_files

### 2018-10-11 document/push.sh
remove quiet

@@ -13,7 +13,7 @@ commit_paper_files() {
 }
 
 upload_files() {
-  git push https://kipkurui:${GITHUB_TOKEN}@github.com/BioinfoNet/hack-paper.git --quiet --all
+  git push https://kipkurui:${GITHUB_TOKEN}@github.com/BioinfoNet/hack-paper.git --set-upstream --all
 }
 setup_git
 commit_paper_files

### 2018-10-11 document/push.sh
checkout temp-branch

@@ -10,6 +10,9 @@ commit_paper_files() {
   git add *.pdf
   git add *.tex
   git commit --message "Travis build: $TRAVIS_BUILD_NUMBER"
+  git branch my-temporary-work
+  git checkout master
+  git merge my-temporary-work
 }
 
 upload_files() {

### 2018-10-11 document/manuscript.tex
Travis build: 32

\section*{Introduction}\label{introduction}
\addcontentsline{toc}{section}{Introduction}

Add an introduction of the Open Science. To cite an article, use
\cite{Schlegel2016}. All the bibliographies should be added to
\texttt{library.bib} in the BibTeX format. See the example in
\texttt{library.bib}.

Since this will more or less be like a review article, we will need to
identify the various subsections based on the topics we need to cover in
the review. See
\href{http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005619}{this
article} for some tips \cite{Mensh2017}.

\subsection*{Status of Open Science in Kenya: Literature
search}\label{status-of-open-science-in-kenya-literature-search}
\addcontentsline{toc}{subsection}{Status of Open Science in Kenya:
Literature search}

The report from the above team will be useful for writing the
introduction, as well as providing materials to be used in the
discussion. In fact, we can use their resource to weave the whole paper
together.

In this section, and the introduction, we will conduct a review of the
status of open science in the country. - What kind of resources are
available to support open science - Are there policies or incentives for
open science practices? - What kind of training activities have been
conducted to promote and train students and researchers on open science
tools and practices? - etc

\section*{DataMining Section}\label{datamining-section}
\addcontentsline{toc}{section}{DataMining Section}

This will be a data analysis section. The title of this section will
depend on the results of your analysis. For example, if Kenyan
researchers are not publishing open access, we will need to understand
why that is the case. The solution may lie in the cost of publishing,
and that is how the resource created by Open Access options team is
useful.

We address questions like: - What is the publishing trend by Kenyan
researchers - Are they publishing open access, and how has this changed
over the years? - Are Kenyan researchers embracing pre-prints (BioRXiv,
AriRXiv, ResearchGate). Who is driving the adoption of pre-prints? Local
researchers or foreign collaborators? - What are the collaboration
trends? Are Kenyan researcher collaborating locally or internationally?

\section*{But Publishing Open Access is
Expensive}\label{but-publishing-open-access-is-expensive}
\addcontentsline{toc}{section}{But Publishing Open Access is Expensive}

This remains the main barrier to publishing open access, in addition to
the obsession with impact factors. However, for early career scientists
and students, especially in developing countries, most publishers offer
waivers and subsidies. In this section, we explore some of the avenues
to publishing open access at low cost.

To address this problem, we created a resource that can guide ECR and
students on where they publish open access, and at low cost. We also
provide information on how they can still be open, eg via the green
route.

\section*{Figures}\label{figures}
\addcontentsline{toc}{section}{Figures}

You can add the figures as follows:

\begin{figure}[htbp]
\centering
%% \includegraphics{./figure01.pdf}
\caption{Figure 1}
\end{figure}

And you can have it referenced as a figure

\begin{quote}
\textbf{Box 1} To highlight of defining some key concepts in Open
science without disrupting the flow of the articles, you can use a quote
format.
\end{quote}

\section*{Discussion}\label{discussion}
\addcontentsline{toc}{section}{Discussion}

What do the results mean? How does your results fit to the current
literature? How do they compare to other similar studies?

\section*{Conclusions}\label{conclusions}
\addcontentsline{toc}{section}{Conclusions}

What is the take-home message from this article?

\section{Acknowledgement}\label{acknowledgement}

We acknowledge the support from \ldots{} and the contribution from
\ldots{}.

Use this section to acknowledge funding and resource contributions to
the project.

### 2018-10-11 document/push.sh
update manuscript

@@ -3,7 +3,6 @@
 setup_git() {
   git config --global user.email "calebkibet88@gmail.com"
   git config --global user.name "Caleb Kibet"
-  
 }
 
 commit_paper_files() {
@@ -18,6 +17,7 @@ commit_paper_files() {
 upload_files() {
   git push https://kipkurui:${GITHUB_TOKEN}@github.com/BioinfoNet/hack-paper.git --set-upstream --all
 }
+
 setup_git
 commit_paper_files
 upload_files

### 2018-11-06 document/push.sh
edit temp branch behaviour

@@ -12,6 +12,7 @@ commit_paper_files() {
   git branch my-temporary-work
   git checkout master
   git merge my-temporary-work
+  git branch -d my-temporary-work
 }
 
 upload_files() {

### 2018-11-06 document/push.sh
Update push.sh

@@ -8,6 +8,7 @@ setup_git() {
 commit_paper_files() {
   git add *.pdf
   git add *.tex
+  git add *.html
   git commit --message "Travis build: $TRAVIS_BUILD_NUMBER"
   git branch my-temporary-work
   git checkout master

### 2018-11-06 document/manuscript.tex
Travis build: 55

@@
 over the years? - Are Kenyan researchers embracing pre-prints (BioRXiv,
-AriRXiv, ResearchGate). Who is driving the adoption of pre-prints? Local
-researchers or foreign collaborators? - What are the collaboration
-trends? Are Kenyan researcher collaborating locally or internationally?
+AriRXiv, ResearchGate, F1000Research). Who is driving the adoption of
+pre-prints? Local researchers or foreign collaborators? - What are the
+collaboration trends? Are Kenyan researcher collaborating locally or
+internationally?
 
@@
 
-This remains the main barrier to publishing open access, in addition to
-the obsession with impact factors. However, for early career scientists
-and students, especially in developing countries, most publishers offer
-waivers and subsidies. In this section, we explore some of the avenues
-to publishing open access at low cost.
+The article processing charge is the main barrier to publishing open
+access, in addition to the obsession with impact factors. However, for
+early career scientists and students, especially in developing
+countries, most publishers offer waivers and subsidies but few are
+aware. In this section, we explore some of the avenues to publishing
+open access at low cost.
 
@@
 students on where they publish open access, and at low cost. We also
-provide information on how they can still be open, eg via the green
-route.
+provide information on how they can still be open when they publsih in
+paywalled journals, eg via the green route.
 
@@
 
-What is the take-home message from this article?
+What is the take-home message from this article? What are the
+recommendations?
 
@@
 We acknowledge the support from \ldots{} and the contribution from
-\ldots{}.
+Mozilla Science Lab and KENET.
 

### 2020-03-09 document/manuscript.tex
Travis build: 67

@@
+\section{Towards Open and Reproducible Genomic Research: Lessons from
+OpenScienceKE}\label{towards-open-and-reproducible-genomic-research-lessons-from-openscienceke}
+
+\subsection*{Abstract}\label{abstract}
+\addcontentsline{toc}{subsection}{Abstract}
+
+Open, collaborative and reproducible research -- Open Science -- has a
+great potential for advancing science. However, the training in our
+local universities does not equip students with the tools to practice
+open science. However, to work in the open and collaborate, your
+collaborators should be equipped to use the tools that you use. The main
+barrier to working open, therefore, is the lack of awareness of the
+collaboration tools and the skills required to utilize these tools.
+Therefore, to fill the gap through an open science community, funded by
+a Mozilla Mini-grant -- OpenScienceKE, we are promoting open science
+among bioinformatics students and researchers in the Nairobi area by
+training using this model: sensitize, train, hack and collaborate. This
+model first sensitizes on open science practices through seminars,
+trains on open science tools through workshops, facilitates hands-on
+application of the tools through hackathons, and finally fosters a
+community of open science enthusiasts through meetups.
+
+OpenScienceKE sought to address the following problems: the lack of
+awareness of open science practices and tools within the Bioinformatics
+community in Kenya; the poor adoption of open science practices in
+Bioinformatics; and the absence of research to establish the state of
+affairs in adopting open science in Kenya. From the OpenScienceKE
+hackathon, we managed to create an open resource that the students could
+use to figure out where they can cost-effectively publish open access.
+In addition, through literature search and data mining, we observed a
+growing interest in open science practices in Kenya but the lack of
+awareness and skills hinder the adoption. The use of preprints for
+research dissemination haven't caught up in Kenya; out of the 20,069
+papers downloaded from bioRXiv, only 18 have Kenyan authors, a majority
+of which are as a result of international (16) collaborations. We also
+observed a lack of incentives and policy in academic and research
+institutions to support open science. The fear of being scooped and the
+competitive spirit within the scientific community are also major
+barriers to working in the open.
+
+The first iteration of the model which focused on academic institutions
+set the foundation for next phase: promote the open and reproducible
+science in research institutions. This model provides the framework for
+the adoption of open science practices within the institution and others
+in the future. As genomic research data generated in Africa grows, there
+is a need for the adoption of open science practices in data storage,
+reproducible pipelines and collaborative research. We propose this
+approach, which develops the necessary infrastructure within research
+institutions, and builds human capacity through the model: sensitize,
+train, hack and collaborate. Promotion of open science in Africa
+recognizes the future direction of research and OpenScienceKE is growing
+the culture and practice in the research active region.
+
 \section*{Introduction}\label{introduction}
 \addcontentsline{toc}{section}{Introduction}
+
+Our Core message is to present a clear outlook on the adoption of Open
+Science practice by African-based scientists, starting with Kenya.
 
@@
 What is the take-home message from this article? What are the
-recommendations?
+recommendations? - The need for a framework to guide the adoption of
+open science practices. Take note of the bariers and provide
+recommendations. - The need for low-cost publishing - The need for
+policies on open science that can be implemented by various
+insituitions. For example, we can provide a template that can be adopted
+by most instituitions
 

### 2020-03-09 document/manuscript.tex
Travis build: 73

@@
-\section{Towards Open and Reproducible Genomic Research: Lessons from
-OpenScienceKE}\label{towards-open-and-reproducible-genomic-research-lessons-from-openscienceke}
+\section*{Introduction}\label{introduction}
+\addcontentsline{toc}{section}{Introduction}
 
-\subsection*{Abstract}\label{abstract}
-\addcontentsline{toc}{subsection}{Abstract}
+Our Core message is to present a clear outlook on the adoption of Open
+Science practice by African-based scientists, starting with Kenya.
+
+Add an introduction of the Open Science. To cite an article, use
+\cite{Schlegel2016}. All the bibliographies should be added to
+\texttt{library.bib} in the BibTeX format. See the example in
+\texttt{library.bib}.
+
+Since this will more or less be like a review article, we will need to
+identify the various subsections based on the topics we need to cover in
+the review. See
+\href{http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005619}{this
+article} for some tips \cite{Mensh2017}.
+
+\textbf{Placeholder Text for the Background}
 
@@
 recognizes the future direction of research and OpenScienceKE is growing
-the culture and practice in the research active region.
-
-\section*{Introduction}\label{introduction}
-\addcontentsline{toc}{section}{Introduction}
-
-Our Core message is to present a clear outlook on the adoption of Open
-Science practice by African-based scientists, starting with Kenya.
-
-Add an introduction of the Open Science. To cite an article, use
-\cite{Schlegel2016}. All the bibliographies should be added to
-\texttt{library.bib} in the BibTeX format. See the example in
-\texttt{library.bib}.
-
-Since this will more or less be like a review article, we will need to
-identify the various subsections based on the topics we need to cover in
-the review. See
-\href{http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005619}{this
-article} for some tips \cite{Mensh2017}.
+the culture and practice in the research acti
 

### 2020-03-09 document/manuscript-biorxiv.tex
Adding pdf file
Change on add remote push.sh

@@
 {\Large
-\textbf\newline{\LaTeX Template for Manuscripts with Embedded Figures.}
+\textbf\newline{Towards Open and Reproducible Genomic Research: Lessons from OpenScienceKE}
 }
@@
 \\
-Author 1\textsuperscript{1},
+Caleb Kibet \textsuperscript{1,3,*},
 Author 2\textsuperscript{2},
@@
 Author 6\textsuperscript{2},
-Author 7\textsuperscript{1,*}
+OpenScienceKE\textsuperscript{3}
 \\
 \bigskip
-\bf{1} Affiliation A
+\bf{1} International Center of Insect Physiology and Ecology
 \\
@@
 \\
+\bf{3} Members of the OpenScienceKE community.
+\\
 \bigskip
-* correseponding@author.mail
+* ckibet@icipe.org
 
@@
 \section*{Abstract}
-Fitting multiple figures into very tight manuscripts while keeping it pleasant to read is challenging.
-Therefore figures are often simply attached to the very end of a manuscript file.
-While easier for the authors, this practice is inconvenient for readers.
-This \LaTeX template shows how to generate a compiled PDF with figures embedded into the text.
-It provides several examples of how to embed figures or tables directly into the text thus giving you a range of options from which you should choose the one best suited for your manuscript.
-Check out Schlegel et al., (2016) as example of use \cite{Schlegel2016}.
+
+Open, collaborative and reproducible research – Open Science – has a great potential for advancing science.
+However, the training in our local universities does not equip students with the tools to practice open science.
+However, to work in the open and collaborate, your collaborators should be equipped to use the tools that you use.
+The main barrier to working open, therefore, is the lack of awareness of the collaboration tools and the skills required to utilize these tools.
+Therefore, to fill the gap through an open science community, funded by a Mozilla Mini-grant – OpenScienceKE, we are promoting open science among bioinformatics students and researchers in the Nairobi area by training using this model: sensitize, train, hack and collaborate.
+This model first sensitizes on open science practices through seminars, trains on open science tools through workshops, facilitates hands-on application of the tools through hackathons, and finally fosters a community of open science enthusiasts through meetups.
+
+OpenScienceKE sought to address the following problems: the lack of awareness of open science practices and tools within the Bioinformatics community in Kenya; the poor adoption of open science practices in Bioinformatics; and the absence of research to establish the state of affairs in adopting open science in Kenya.
+From the OpenScienceKE hackathon, we managed to create an open resource that the students could use to figure out where they can cost-effectively publish open access.
+In addition, through literature search and data mining, we observed a growing interest in open science practices in Kenya but the lack of awareness and skills hinder the adoption.
+The use of preprints for research dissemination haven’t caught up in Kenya; out of the 20,069 papers downloaded from bioRXiv, only 18 have Kenyan authors, a majority of which are as a result of international (16) collaborations.
+We also observed a lack of incentives and policy in academic and research institutions to support open science.
+The fear of being scooped and the competitive spirit within the scientific community are also major barriers to working in the open.
+
+The first iteration of the model which focused on academic institutions set the foundation for next phase: promote the open and reproducible science in research institutions.
+This model provides the framework for the adoption of open science practices within the institution and others in the future.
+As genomic research data generated in Africa grows, there is a need for the adoption of open science practices in data storage, reproducible pipelines and collaborative research.
+We propose this approach, which develops the necessary infrastructure within research institutions, and builds human capacity through the model: sensitize, train, hack and collaborate.
+Promotion of open science in Africa recognizes the future direction of research and OpenScienceKE is growing the culture and practice in the research active region.
 

### 2020-03-10 document/manuscript-out.tex
Travis build: 125

--- 2020-03-09 document/manuscript.tex
@@
 
+\subsubsection{Proposed subsections and
+content}\label{proposed-subsections-and-content}
+
+The subsections will be based on the findings about the status of open
+science in Kenya.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Resources to support open science are insufficient. The resources
+  include:
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  ICT infrastructure. KENET provides infrastructure for open science
+  training.
+\item
+  Institutional repositories. Universities and Research institutes have
+  repositories to store publications. Also repositories to store
+  research data. Do all institutions have these repositories?
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Lacking policies and incentives from the government, local funders and
+  institutions. These policies and incentives lack in open science
+  practices such as:
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  Research data management. Apart from
+  \href{https://cgspace.cgiar.org/handle/10568/63496}{ILRI's data
+  management policy}, do other institutions have these policies?
+\item
+  Open access publishing
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Limited training on Open Science. Awareness on Open Science has
+  previously been through:
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  Open access week in universities. (Do research institutions
+  participate in Open Access week?)
+\item
+  Workshops such as OpenScienceKE Workshop in 2018
+\item
+  Others?
+\end{itemize}
+
 \section*{DataMining Section}\label{datamining-section}
@@
 
+\subsubsection{Proposed title and
+content}\label{proposed-title-and-content}
+
+The title here should indicate whether \emph{Kenyan researchers are
+embracing open science practices} as per the findings of the Data Mining
+team. This section's discussion may therefore focus on supporting this
+claim based on:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  The publishing trend of Kenyan researchers. The trend indicates that
+  open access publishing is popular. The popular journals where Kenyan
+  researcher publish are open access.
+\item
+  Pre-prints. \textbf{\emph{more work needed here to find out whether
+  they have embraced preprints and if so, who is driving the adoption?}}
+\item
+  Open data and Code.
+  \textbf{\emph{\href{https://github.com/BioinfoNet/Data-mining/issues/11}{This
+  is yet to be completed}. We need to find out whether Kenyan authors
+  are making the data and code available and accessible, and whether
+  they are adhering to the FAIR principles}}
+\item
+  Collaboration trends. Are Kenyan authors collaborating from within or
+  without? This milestone is still pending. 5 What else?
+\end{enumerate}
+
 \section*{But Publishing Open Access is
@@
 paywalled journals, eg via the green route.
+
+\subsubsection{Proposed content of
+Section}\label{proposed-content-of-section}
+
+\begin{itemize}
+\tightlist
+\item
+  Low cost open access publishing is available.
+\end{itemize}
+
+Here our discussion will indicate that there are publishers offering
+waivers and subsidies for developing countries (low income and lower
+middle income countries). Waiver is either Full or 50\% of the APCs.
+Sometimes, waivers are offered on a case-by-case basis. The resource
+created will create awareness on where to publish at low cost in an open
+access way.
 

### 2020-03-11 document/manuscript-out.tex
Travis build: 128
Travis build: 129

@@
 
-\textbf{Placeholder Text for the Background}
+\textbf{Placeholder Text for the Background} WILL UPDATE TO INTRODUCTION
 
@@
 
-\section*{DataMining Section}\label{datamining-section}
-\addcontentsline{toc}{section}{DataMining Section}
+\section*{Data Mining Section}\label{data-mining-section}
+\addcontentsline{toc}{section}{Data Mining Section}
 
@@
 \item
-  Pre-prints. \textbf{\emph{more work needed here to find out whether
-  they have embraced preprints and if so, who is driving the adoption?}}
+  Pre-prints. Pre-prints are yet to be widely adopted by Kenyan
+  scientists and foreign collaborators drive the adoption of pre-prints
+  among Kenyan authors
 \item
@@
   Collaboration trends. Are Kenyan authors collaborating from within or
-  without? This milestone is still pending. 5 What else?
+  without? (based on pre-prints from bioRXiv, we see that Kenyan authors
+  are mostly collaborating internationally. Is this the case from the
+  rest of literature obtained during data mining?) 5 What else?
 \end{enumerate}
@@
 students on where they publish open access, and at low cost. We also
-provide information on how they can still be open when they publsih in
+provide information on how they can still be open when they publish in
 paywalled journals, eg via the green route.
@@
 recommendations? - The need for a framework to guide the adoption of
-open science practices. Take note of the bariers and provide
+open science practices. Take note of the barriers and provide
 recommendations. - The need for low-cost publishing - The need for
 policies on open science that can be implemented by various
-insituitions. For example, we can provide a template that can be adopted
-by most instituitions
+institutions. For example, we can provide a template that can be adopted
+by most institutions
 
@@
 We acknowledge the support from \ldots{} and the contribution from
-Mozilla Science Lab and KENET.
+Mozilla Science Lab and KENET, and ICIPE.
 

### 2020-03-29 document/manuscript-out.tex
Travis build: 130

@@
 
+\subsubsection{Draft text for the
+section}\label{draft-text-for-the-section}
+
+Access to literature is of paramount importance in the process of
+scientific inquiry. The free flow of ideas and information are crucial
+to addressing economic, environmental and social issues facing
+developing countries especially in Africa. However, contrary to
+concerted research efforts to foster development, an assessment of
+Africa's research output between 1999 - 2008 indicated that Africa
+lagged behind in it's research output when compared to the rest of the
+world \cite{adams2010;raju2012}. Poor accessibility of scientific
+publications has been suggested as one of the reasons for the low
+research output from Africa \cite{tise2011;raju2012}.
+
 \section*{Figures}\label{figures}

### 2020-07-12 document/manuscript-out.tex
Travis build: 131

@@
 publications has been suggested as one of the reasons for the low
-research output from Africa \cite{tise2011;raju2012}.
+research output from Africa \cite{tise2011;raju2012}. Accessibility of
+scientific literature by African researchers is mostly hindered by the
+lack of access to subscription-based journal articles due to their
+unaffordable cost. However, the main barrier to equitable access and
+sharing of knowledge in Africa is the cost publishing, which is
+unaffordable for most researchers in Low- and Middle-income countries.
+
+The concept of openness in scientific research has long been proposed as
+a means toward equitable sharing of knowledge
+\cite{burton_sharing_2011;nikos_building_2019}. Consequently, most
+journals have shifted from subscription-based access to research
+articles and embraced open access publishing. In this case, however, the
+authors have to pay article publication fees which often run into
+thousands of dollars - another barrier to publishing especially from
+authors in Low- and Middle-income countries. In most Kenyan
+institutions, for example, students have to publish at least one paper
+at MSc level. With limited funds and the need to publish in a reputable
+journal, most students usually don't know which journals to publish in
+due to the high cost of article processing charges (APCs).
+
+In this study, we observed that Kenyan authors are increasingly adopting
+open access publishing, a trend that has been observed in Low Income
+countries \cite{iyandemye_low_2019}. To help students and early career
+scientists adopt this open access publishing trend, we looked into
+publishers that publish open access and created a resource that acts as
+a guideline for them to publish open access, and at low cost. We found
+that most Open Access publishers offer waivers or subsidies to authors
+from developing countries. The publishers had different criteria for
+offering waivers. Some publishers, such as Nature Publishing Group,
+offered a full waiver to authors in low-income countries and 50\%
+discount for authors in low-middle-income countries. Some of the
+publishers' waiver policies are linked to the corresponding author,
+source of funds for research or the financial need of the author. For
+example, PLOS offers waivers of about 50\% or more if the research was
+carried out in low or LMIC countries therefore linking their waiver
+policy to source of research funds. Publishers such as Oxford University
+Press and eLIFE offer waivers to authors in low- and middle-income
+countries who experience financial hardships or have insufficient funds.
+Waivers are also available on a case-by-case basis for authors in low-
+and middle-income countries by publishers such as MDPI.
+
+As a low-middle-income country, Kenyan authors are eligible for waivers
+and subsidies. Although we found no special waiver to students
+identified so far, students may apply for waivers as authors in
+financial need in developing countries whom publishers offer waivers on
+a case-by-case basis. Additionally, students may also place their
+manuscript on preprint servers, the green route of publishing, where
+they can receive feedback on their work before publication in an open
+access journal. Both the use of preprint servers and open access
+publishing by taking advantage of the waivers offered by publishers will
+increase the accessibility of research literature in Africa.
 
